---
title: "2_PP_active_retire"
output: html_document
date created: 1/5/23
date updated: 5/26/23

Purpose: Explore whether the proportion of potentially vulnerable groups within the buffers surrounding power plants that are still active differ from those that have been retired since 2010.
---

Clear environment.
```{r}
rm(list=ls())
gc()

```

=========================================================================================================
Load packages.
```{r, include=FALSE}
library(foreign)
library(tidyverse)
library(tidycensus)
library(cowplot)
library(spatstat)

library(sf)
library(terra)
library(tmap)

```

Load functions - change to your directory.
```{r}
source('f_Misc.R')
source('f_census.R')

```

=========================================================================================================
Load data - change to your directory.
--------------------------------------

Spatial data.
```{r}
## US states
usa_shp0 <- st_read('data/States_shapefile.shp')
# Remove AK, HI
usa_shp <- filter(usa_shp0, !(State_Code %in% c('AK', 'HI')))
rm(usa_shp0)

## EGU locations
egu_pt0 <- st_read('data/facility2020_pt.shp')
egu_pt  <- st_transform(egu_pt0, st_crs(usa_shp))
rm(egu_pt0)

## Census bg shp
bg20_shp0 <- st_read('data/US_blck_grp_2020.shp')
# Transform crs
bg20_shp <- st_transform(bg20_shp0, st_crs(usa_shp))
rm(bg20_shp0)

```

Load MATS data.
```{r}
## Boiler
# boiler_change <- read.csv('/Users/monadai/OneDrive - Harvard University/Harvard/Sunderland/MATS/Data/FromBen/boiler_level_change_2010_2020.csv')
# head(boiler_change)

```

Load buffer data.
```{r}
# Facility ID & census BG.
egu_bg_0  <- read.dbf('data/facility_gisjoin5.dbf')
egu2020_0 <- read.csv('data/facility2020_retire_0123.csv')

# Load US shp.
usa_shp0 <- st_read('data/States_shapefile.shp')

```

====================================================================================================================
Calculate proportion of subgroup per census block group.
------------------------------------------------------------------

Collect correct census variables from NHGIS dataframe (see census2020.R).
```{r}
source('data/census2020.R')

```

Create new df from loaded US census data combining estimate & MOE data.
```{r}
# FISH
fish_df00 <- LoadFISH(nhgisYR = nhgis2020_bg, nhgiscode_est = nhgis2020_EST, nhgiscode_moe = nhgis2020_MOE)
fish_df0  <- LoadEJdf(fish_df00)

fish_df0  <- fish_df0 %>% group_by(GISJOIN) %>% mutate(FISH = sum(AIAN, ASIAN, HIPI, MULTI))

# --------------------------------------------------------------------------------------------------------------------
# INCOME
income_df00 <- LoadINCOME(nhgisYR = nhgis2020_bg, nhgiscode_est = nhgis2020_EST, nhgiscode_moe = nhgis2020_MOE)
income_df0  <- LoadEJdf(income_df00)

income_df0  <- income_df0 %>% group_by(GISJOIN) %>% mutate(INCOME20 = sum(L10k, L15k, L20k))
# head(income_df0)

# --------------------------------------------------------------------------------------------------------------------
# POVERTY
poverty_df00 <- LoadPOVERTY(nhgisYR = nhgis2020_bg, nhgiscode_est = nhgis2020_EST, nhgiscode_moe = nhgis2020_MOE)
poverty_df0  <- LoadEJdf(poverty_df00)

poverty_df0  <- poverty_df0 %>% group_by(GISJOIN) %>% 
  mutate(POVERTY200 = sum(R0.50, R1.00, R1.25, R1.50, R1.85, R2.00))
# head(poverty_df0)

# --------------------------------------------------------------------------------------------------------------------
# NOENG
noeng_df00 <- LoadNOENG(nhgisYR = nhgis2020_bg, nhgiscode_est = nhgis2020_EST, nhgiscode_moe = nhgis2020_MOE)
noeng_df0  <- LoadEJdf(noeng_df00)

noeng_df0  <- noeng_df0 %>% group_by(GISJOIN) %>% 
  mutate(NOENG = sum(Sp, Euro, Azn, Other))
# head(noeng_df0)

# --------------------------------------------------------------------------------------------------------------------
# NOHISCH
nohisch_df00 <- LoadNOHISCH(nhgisYR = nhgis2020_bg, nhgiscode_est = nhgis2020_EST, nhgiscode_moe = nhgis2020_MOE)
nohisch_df0  <- LoadEJdf(nohisch_df00)

nohisch_df0  <- nohisch_df0 %>% group_by(GISJOIN) %>% 
  mutate(NOHISCH = sum(edu1, edu2, edu3, edu4, edu5,
                     edu6, edu7, edu8, edu9,
                     edu10, edu11, edu12, edu13, 
                     edu14, edu15))
# head(nohisch_df0)
```

Calculate MOE.
```{r}
# FISH
fish_df <- CalcMOE(fish_df0, 'FISH', 'RACE_Tot')
# head(fish_df)
# summary(fish_df)

income_df <- CalcMOE(income_df0, 'INCOME20', 'INCOME_Tot')
# head(income_df)
# summary(income_df)

poverty_df <- CalcMOE(poverty_df0, 'POVERTY200', 'POVERTY_Tot')
# head(poverty_df)
# summary(poverty_df)

noeng_df <- CalcMOE(noeng_df0, 'NOENG', 'NOENG_Tot')
# head(noeng_df)
# summary(noeng_df)

nohisch_df <- CalcMOE(nohisch_df0, 'NOHISCH', 'NOHISCH_Tot')
# head(nohisch_df)
# summary(nohisch_df)
```

Join all EJ factors into 1 df.
```{r}
vulnerable20_tot <- left_join(fish_df0[c(1,2,4, ncol(fish_df0))], income_df0[,c(1,4, ncol(income_df0))], by = 'GISJOIN')
vulnerable20_tot <- left_join(vulnerable20_tot, poverty_df0[,c(1,4, ncol(poverty_df0))], by = 'GISJOIN')
vulnerable20_tot <- left_join(vulnerable20_tot, noeng_df0[,c(1,4, ncol(noeng_df0))], by = 'GISJOIN')
vulnerable20_tot <- left_join(vulnerable20_tot, nohisch_df0[,c(1,4, ncol(nohisch_df0))], by = 'GISJOIN')

head(vulnerable20_tot)

```

=========================================================================================================
Clean data.
```{r}
# Separate col into facility ID & census BG
egu_bg <- separate(egu_bg_0, facility_g, c('facility_id_master', 'GISJOIN'))
# Join census BG data
egu_bg <- left_join(egu_bg, vulnerable20_tot, by = 'GISJOIN')
# Take out Puerto Rico facilities
egu_bg <- filter(egu_bg, !(facility_id_master %in% c(68029, 70000, 70001, 70002, 70003)))
egu_bg

## Save 512 unique power plants.
egu2020 <- egu2020_0 %>% dplyr::select('facility_id_master', 'Facility.Retire') %>% unique()
# Take out Puerto Rico facilities
egu2020 <- filter(egu2020, !(facility_id_master %in% c(68029, 70000, 70001, 70002, 70003)))
egu2020$facility_id_master <- as.character(egu2020$facility_id_master)
egu2020

```

=========================================================================================================
Separate retired & active plants.

```{r}
# Combine census BG w/ facility status
egu_ppl0 <- left_join(egu2020, egu_bg, by = "facility_id_master")
egu_ppl1 <- egu_ppl0 %>% 
  mutate(FISHppl    = FISH/RACE_Tot * POP,
         POVERTYppl = POVERTY200/POVERTY_Tot * POP,
         INCOMEppl  = INCOME20/INCOME_Tot * POP,
         NOHISCHppl = NOHISCH/NOHISCH_Tot * POP,
         NOENGppl   = NOENG/NOENG_Tot * POP)
egu_ppl1

egu_ppl2 <- egu_ppl1 %>% group_by(facility_id_master) %>%
  mutate(FISHpct    = sum(FISHppl, na.rm = T) / sum(POP, na.rm = T, na.rm = T),
         POVERTYpct = sum(POVERTYppl, na.rm = T) / sum(POP, na.rm = T, na.rm = T),
         INCOMEpct  = sum(INCOMEppl, na.rm = T) / sum(POP, na.rm = T, na.rm = T),
         NOHISCHpct = sum(NOHISCHppl, na.rm = T) / sum(POP, na.rm = T, na.rm = T),
         NOENGpct   = sum(NOENGppl, na.rm = T) / sum(POP, na.rm = T, na.rm = T),
               
         POPppl     = POP %>% sum(na.rm = T)
  )

egu_ppl <- egu_ppl2  %>% 
  distinct(facility_id_master, Facility.Retire, POPppl, 
           FISHpct, POVERTYpct, INCOMEpct, NOHISCHpct, NOENGpct) %>%
  dplyr::select(facility_id_master, Facility.Retire, POPppl, 
                FISHpct, POVERTYpct, INCOMEpct, NOHISCHpct, NOENGpct)
head(egu_ppl)

# Split into active & retired plants
egu_active <- filter(egu_ppl, Facility.Retire != "Fully.retired") %>% data.frame()
egu_retire <- filter(egu_ppl, Facility.Retire == "Fully.retired") %>% data.frame()

egu_active %>% summary()
egu_retire %>% summary()

```

Perform Wilcoxon test.
```{r}
## Find weighted medians
# Create empty df
wt_med <- data.frame(matrix(nrow = 5, ncol = 2))
wt_quant95_active <- data.frame(matrix(nrow = 5, ncol = 2))
wt_quant95_retire <- data.frame(matrix(nrow = 5, ncol = 2))
# Active
for (i in 1:5){
  wt_med[i,1] <- weighted.median(egu_active[, 3+i], w = egu_active[, 'POPppl'], na.rm = T)

  wt_quant95_active[i,1] <- weighted.quantile(egu_active[, 3+i], w = egu_active[, 'POPppl'], na.rm = T, probs=0.025)
  wt_quant95_active[i,2] <- weighted.quantile(egu_active[, 3+i], w = egu_active[, 'POPppl'], na.rm = T, probs=0.975)
  rownames(wt_quant95_active) <- c('FISH', 'POVERTY', 'INCOME', 'NOHISCH', 'NOENG')
}

# Retired
for (i in 1:5){
  wt_med[i,2] <- weighted.median(egu_retire[, 3+i], w = egu_retire[, 'POPppl'], na.rm = T)
  
  wt_quant95_retire[i,1] <- weighted.quantile(egu_retire[, 3+i], w = egu_retire[, 'POPppl'], na.rm = T, probs=0.025)
  wt_quant95_retire[i,2] <- weighted.quantile(egu_retire[, 3+i], w = egu_retire[, 'POPppl'], na.rm = T, probs=0.975)
  rownames(wt_quant95_retire) <- c('FISH', 'POVERTY', 'INCOME', 'NOHISCH', 'NOENG')
}
# Add colnames
colnames(wt_med) <- c('Active', 'Retired')


# wilcox test
AvsR_ls <- list(
  fish    = wilcox.test(egu_active$FISHpct,    egu_retire$FISHpct,    conf.int = T),
  poverty = wilcox.test(egu_active$POVERTYpct, egu_retire$POVERTYpct, conf.int = T),
  income  = wilcox.test(egu_active$INCOMEpct,  egu_retire$INCOMEpct,  conf.int = T),
  nohisch = wilcox.test(egu_active$NOHISCHpct, egu_retire$NOHISCHpct, conf.int = T),
  noeng   = wilcox.test(egu_active$NOENGpct,   egu_retire$NOENGpct,   conf.int = T)
)

AvsR_ci <- AvsR_pval <- data.frame(
  row.names = c('FISH', 'POVERTY', 'INCOME', 'NOHISCH', 'NOENG'),
  Low = c(
    AvsR_ls$fish$conf.int[1],
    AvsR_ls$poverty$conf.int[1],
    AvsR_ls$income$conf.int[1],
    AvsR_ls$nohisch$conf.int[1],
    AvsR_ls$noeng$conf.int[1]
  ),
  High = c(
    AvsR_ls$fish$conf.int[2],
    AvsR_ls$poverty$conf.int[2],
    AvsR_ls$income$conf.int[2],
    AvsR_ls$nohisch$conf.int[2],
    AvsR_ls$noeng$conf.int[2]
  )
) 
AvsR_ci

AvsR_pval <- data.frame(FISH    = AvsR_ls$fish$p.value,
                        POVERTY = AvsR_ls$poverty$p.value,
                        INCOME  = AvsR_ls$income$p.value,
                        NOHISCH = AvsR_ls$nohisch$p.value,
                        NOENG   = AvsR_ls$noeng$p.value) %>% t()
colnames(AvsR_pval) <- 'pval'

AvsR <- cbind(wt_med, AvsR_ci, AvsR_pval)
KnitTab(AvsR)

```


=========================================================================================================
Look at change in subpopulations near active powerplants over multiple buffer sizes.
```{r}
# Take out duplicates
egu_pt1 <- egu_pt[which(duplicated(egu_pt$facility_i) == F), ]

# Active plants only
active_pt     <- filter(egu_pt1, Facility_R != 'Fully.retired')
# Transform EGU locations into crs with units in [m]
active_pt7801 <- st_transform(active_pt, crs = 7801)
active_pt7801 <- active_pt7801[which(duplicated(active_pt7801$facility_i) == F), ]
head(active_pt7801)

# Retire plants only
retire_pt     <- filter(egu_pt1, Facility_R == 'Fully.retired')
# Transform EGU locations into crs with units in [m]
retire_pt7801 <- st_transform(retire_pt, crs = 7801)
retire_pt7801 <- retire_pt7801[which(duplicated(retire_pt7801$facility_i) == F), ]
head(retire_pt7801)

```

Perform buffer.
```{r}
sf_use_s2(FALSE)

buff1   <- Compare_EGU_Buffers(buff.sz = "1", us_egu = active_pt7801, bg_shp = bg20_shp)
buff1.5 <- Compare_EGU_Buffers(buff.sz = "1.5", us_egu = active_pt7801, bg_shp = bg20_shp)
buff2   <- Compare_EGU_Buffers(buff.sz = "2", us_egu = active_pt7801, bg_shp = bg20_shp)
buff3   <- Compare_EGU_Buffers(buff.sz = "3", us_egu = active_pt7801, bg_shp = bg20_shp)
buff4   <- Compare_EGU_Buffers(buff.sz = "4", us_egu = active_pt7801, bg_shp = bg20_shp)
buff5   <- Compare_EGU_Buffers(buff.sz = "5", us_egu = active_pt7801, bg_shp = bg20_shp)
buff10  <- Compare_EGU_Buffers(buff.sz = "10", us_egu = active_pt7801, bg_shp = bg20_shp)
buff15  <- Compare_EGU_Buffers(buff.sz = "15", us_egu = active_pt7801, bg_shp = bg20_shp)
buff20  <- Compare_EGU_Buffers(buff.sz = "20", us_egu = active_pt7801, bg_shp = bg20_shp)
buff25  <- Compare_EGU_Buffers(buff.sz = "25", us_egu = active_pt7801, bg_shp = bg20_shp)
buff50  <- Compare_EGU_Buffers(buff.sz = "50", us_egu = active_pt7801, bg_shp = bg20_shp)
buff75  <- Compare_EGU_Buffers(buff.sz = "75", us_egu = active_pt7801, bg_shp = bg20_shp)
buff100 <- Compare_EGU_Buffers(buff.sz = "100", us_egu = active_pt7801, bg_shp = bg20_shp)

allbuff <- rbind(buff1, buff1.5, buff2, buff3, buff4, buff5, buff10, buff15, buff20, buff25, buff50, buff75, buff100)
allbuff
```

Plot.
```{r}
## Entire US
a <- PlotBuff('FISH', ej_df = fish_df, allbuff = allbuff) + lims(y = c(0.06, 0.115)) + labs(title = 'High-Frequency Fish \nConsumers')
b <- PlotBuff('INCOME20', ej_df = income_df, allbuff = allbuff) + lims(y = c(0.14, 0.185)) + labs(title = '<$20K USD Income')
c <- PlotBuff('POVERTY200', ej_df = poverty_df, allbuff = allbuff) + lims(y = c(0.285, 0.365)) + labs(title = 'PIR<2')
d <- PlotBuff('NOHISCH', ej_df = nohisch_df, allbuff = allbuff) + lims(y = c(0.10, 0.14)) + labs(title = '<High School')
e <- PlotBuff('NOENG', ej_df = noeng_df, allbuff = allbuff) + lims(y = c(0.01, 0.05)) + labs(title = 'Limited English-Speaking')

a
b
c
d
e

plot_grid(a,b,c,
          d,e)

```

==============
For Table 1 in paper, data concerning 5km buffers around power plants.
```{r}
active_buff5   <- Compare_EGU_Buffers(buff.sz = "5", us_egu = active_pt7801, bg_shp = bg20_shp)
retire_buff5   <- Compare_EGU_Buffers(buff.sz = "5", us_egu = retire_pt7801, bg_shp = bg20_shp)

active_buff5
retire_buff5

```
