---
title: '3_Hg_deposition'
date created: 1/11/23
date updated: 5/26/23
Purpose: Evaluate spatial clustering of Hg deposition & sociodemographic factors across the US.
---
Clear environment.
```{r}
rm(list=ls())
gc()

```

=========================================================================================================
Load libraries.
--------------------
```{r, include = F}
library(tidyverse)
library(rsample) 
library(cowplot)

library(spatstat)
library(lmtest)

# Spatial packages
library(raster)
library(ncdf4)
library(sf)
library(tmap)
library(stars)

library(spatialreg)
library(spdep)

```

---------------------------------------------------------------------------------------------------------
Functions - change to your directory.
----------
Spatial model functions.
```{r}
source('f_spatial.R')

```

---------------------------------------------------------------------------------------------------------
Load data - change to your directory.
----------------------------------------
Deposition data.
```{r}
total_deposit2010_ras = stack("data/Total_2010.nc", varname = "Total_Deposition")
total_deposit2020_ras = stack("data/Total_2020.nc", varname = "Total_Deposition")

egu_deposit2010_ras = stack("data/EGU_2010.nc", varname = "EGU_total")
egu_deposit2020_ras = stack("data/EGU_2020.nc", varname = "EGU_total")

```

Load US shp data.
```{r}
usa_shp0 <- st_read('data/States_shapefile.shp')
us_vulnerable_shp0 <- st_read("data/us_vulnerable2020.shp")

# Load 2020 vulnerable subgroups.
vulnerable20_tot <- readRDS('data/vulnerable20_tot.RDS')

```

=========================================================================================================
Create deposition rasters
--------------------
2010 deposition.
```{r}
# Find EGU attributable deposition
egu_att_deposit2010_ras <- egu_deposit2010_ras / total_deposit2010_ras
# Check range of attributable deposition
cellStats(egu_att_deposit2010_ras, stat = 'range')
cellStats(egu_att_deposit2010_ras, stat = 'mean')

# Plot
plot(egu_deposit2010_ras)
plot(total_deposit2010_ras)
plot(egu_att_deposit2010_ras)

```

2020 deposition.
```{r}
# Find EGU attributable deposition
egu_att_deposit2020_ras <- egu_deposit2020_ras / total_deposit2020_ras
# Check range of attributable deposition
cellStats(egu_att_deposit2020_ras, stat = 'range')
cellStats(egu_att_deposit2020_ras, stat = 'mean')

# Plot
plot(egu_deposit2020_ras)
plot(total_deposit2020_ras)
plot(egu_att_deposit2020_ras)

```

==================================================================================================================
Find change in deposition across US between 2010-2020.
---------------------------------------------------------------------------------------------------------
Save contiguous USA data to transform other shp files.
```{r}
# Remove AK, HI
usa_shp <- filter(usa_shp0, !(State_Code %in% c('AK', 'HI')))
# Change crs for us_vulnerable_shp
us_vulnerable_shp <- st_transform(us_vulnerable_shp0, crs = st_crs(usa_shp))

## Replace proportions with number of ppl.
us_vulnerable_shp <- left_join(us_vulnerable_shp[,c('GISJOIN', 'STATE')], vulnerable20_tot, by = "GISJOIN")
head(us_vulnerable_shp)

```

Take out places where deposition = 0 in 2010 (can't reduce deposition if already at 0.
```{r}
dep10_ras <- egu_att_deposit2010_ras
dep20_ras <- egu_att_deposit2020_ras 

dep20_ras[dep10_ras[] == 0] = NA
dep10_ras[dep10_ras[] == 0] = NA

table(is.na(dep20_ras[]))
table(is.na(dep10_ras[]))

dep10_ras %>% plot()
dep20_ras %>% plot()

plot(dep10_ras, colNA="red")
plot(dep20_ras, colNA="red")

```

Find change in deposition between 2010 & 2020.
```{r}
# Find difference in magnitude wrt 2010 deposition
delta_deposit_ras0 <- (dep10_ras - dep20_ras)/dep10_ras
# delta_deposit_ras0 <- (egu_direct_deposition_ras$X2010 - egu_direct_deposition_ras$X2020)/egu_direct_deposition_ras$X2010
delta_deposit_ras  <- delta_deposit_ras0

# Change coordinate ref system (crs) to plot with US
crs(delta_deposit_ras) <- crs(usa_shp)

# Get rid of extra values outside of contiguous US
delta_deposit_mask <- mask(delta_deposit_ras, usa_shp)
delta_deposit_mask2 <- delta_deposit_mask
# delta_deposit_mask2[delta_deposit_mask2[] >0.9] = NA

# Plot
crop(delta_deposit_mask2, usa_shp) %>% plot()
plot(usa_shp %>% st_geometry(), add = T)

```

==================================================================================================================
Perform spatial autocorrelation - (see http://www.geo.hunter.cuny.edu/~ssun/R-Spatial/spregression.html#spatial-error-and-lag-models)
```{r}
## Turn delta_deposit_mask from raster into polygons
delta_deposit_poly <- sf::as_Spatial(sf::st_as_sf(stars::st_as_stars(delta_deposit_mask), 
                            as_points = FALSE, merge = TRUE))
delta_deposit_sf   <- delta_deposit_poly %>% st_as_sf() %>% st_transform(st_crs(usa_shp))

## Join deposition data with 2020 census BG data
sf_use_s2(FALSE)
# delta_vulnerable_shp <- st_join(us_vulnerable_shp, delta_deposit_sf)
delta_vulnerable_shp <- st_join(delta_deposit_sf, us_vulnerable_shp)
colnames(delta_vulnerable_shp)[1] <- 'deltaHg'

head(delta_vulnerable_shp)

```

Calculate proportion of subgroups within each raster cell.
```{r}
## Calculate proportions.
delta_vulnerable_shp11 <- delta_vulnerable_shp %>% 
  mutate(FISHppl    = FISH/RACE_Tot * POP,
         POVERTYppl = POVERTY200/POVERTY_Tot * POP,
         INCOMEppl  = INCOME20/INCOME_Tot * POP,
         NOHISCHppl = NOHISCH/NOHISCH_Tot * POP,
         NOENGppl   = NOENG/NOENG_Tot * POP)
delta_vulnerable_shp11

delta_vulnerable_shp1 <- delta_vulnerable_shp11 %>% group_by(deltaHg) %>%
  mutate(FISH       = sum(FISHppl, na.rm = T) / sum(POP, na.rm = T, na.rm = T),
         POVERTY    = sum(POVERTYppl, na.rm = T) / sum(POP, na.rm = T, na.rm = T),
         INCOME20   = sum(INCOMEppl, na.rm = T) / sum(POP, na.rm = T, na.rm = T),
         NOHISCH    = sum(NOHISCHppl, na.rm = T) / sum(POP, na.rm = T, na.rm = T),
         NOENG      = sum(NOENGppl, na.rm = T) / sum(POP, na.rm = T, na.rm = T),
               
         POPppl     = POP %>% sum(na.rm = T)
  )
```

Data summary/cleaning for subgroups per raster cell.
```{r}
## Range of values
range(delta_vulnerable_shp1$FISH, na.rm = T) 
range(delta_vulnerable_shp1$POVERTY, na.rm = T)
range(delta_vulnerable_shp1$INCOME20, na.rm = T)
range(delta_vulnerable_shp1$NOHISCH, na.rm = T)
range(delta_vulnerable_shp1$NOENG, na.rm = T)

## How many NAs?
filter(delta_vulnerable_shp1, is.na(FISH) == T) %>% nrow()
filter(delta_vulnerable_shp1, is.na(POVERTY) == T) %>% nrow()
filter(delta_vulnerable_shp1, is.na(INCOME20) == T) %>% nrow()
filter(delta_vulnerable_shp1, is.na(NOHISCH) == T) %>% nrow()
filter(delta_vulnerable_shp1, is.na(NOENG) == T) %>% nrow()

## Save only relevant EJ factors.
delta_vulnerable_shp2 <- delta_vulnerable_shp1 %>% dplyr::select('deltaHg', 'STATE', 
'FISH', 'INCOME20', 'POVERTY', 'NOHISCH', 'NOENG')
head(delta_vulnerable_shp2)

## Remove duplicated values.
delta_vulnerable_shp2 <- delta_vulnerable_shp2 %>% unique() %>% na.omit()

## Save as plots.
a <- ggplot(data = delta_vulnerable_shp2) +
  geom_sf(aes(fill = deltaHg), color = NA) + geom_sf(data = usa_shp, alpha = 0, color = 'white') + theme_bw()
b <- ggplot(data = delta_vulnerable_shp2) +
  geom_sf(aes(fill = FISH), color = NA) + geom_sf(data = usa_shp, alpha = 0, color = 'white') + theme_bw()
c <- ggplot(data = delta_vulnerable_shp2) +
  geom_sf(aes(fill = INCOME20), color = NA) + geom_sf(data = usa_shp, alpha = 0, color = 'white') + theme_bw()
d <- ggplot(data = delta_vulnerable_shp2) +
  geom_sf(aes(fill = NOHISCH), color = NA) + geom_sf(data = usa_shp, alpha = 0, color = 'white') + theme_bw()
e <- ggplot(data = delta_vulnerable_shp2) +
  geom_sf(aes(fill = NOENG), color = NA) + geom_sf(data = usa_shp, alpha = 0, color = 'white') + theme_bw()

a
b
c
d
e

# plot_grid(a,b,c,d,e)

```

Impute values.
```{r}
delta_vulnerable_shp3 <- delta_vulnerable_shp2

min_FISH     <- filter(delta_vulnerable_shp3, FISH > 0)$FISH %>% min(na.rm = T)
min_POVERTY  <- filter(delta_vulnerable_shp3, POVERTY > 0)$POVERTY %>% min(na.rm = T)
min_INCOME20 <- filter(delta_vulnerable_shp3, INCOME20 > 0)$INCOME20 %>% min(na.rm = T)
min_NOENG    <- filter(delta_vulnerable_shp3, NOENG > 0)$NOENG %>% min(na.rm = T)
min_NOHISCH  <- filter(delta_vulnerable_shp3, NOHISCH > 0)$NOHISCH %>% min(na.rm = T)

delta_vulnerable_shp3[which(delta_vulnerable_shp3$FISH == 0), 'FISH']         <- min_FISH/2
delta_vulnerable_shp3[which(delta_vulnerable_shp3$POVERTY == 0), 'POVERTY']   <- min_POVERTY/2
delta_vulnerable_shp3[which(delta_vulnerable_shp3$INCOME20 == 0), 'INCOME20'] <- min_INCOME20/2
delta_vulnerable_shp3[which(delta_vulnerable_shp3$NOENG == 0), 'NOENG']       <- min_NOENG/2
delta_vulnerable_shp3[which(delta_vulnerable_shp3$NOHISCH == 0), 'NOHISCH']   <- min_NOHISCH/2

```

Plot histograms to see how many rows are imputed & assess distribution.
```{r}
# Raw data
ggplot(delta_vulnerable_shp3) + geom_histogram(aes(x=deltaHg),  bins=50)
ggplot(delta_vulnerable_shp3) + geom_histogram(aes(x=POVERTY),  bins=50)
ggplot(delta_vulnerable_shp3) + geom_histogram(aes(x=FISH),     bins=50)
ggplot(delta_vulnerable_shp3) + geom_histogram(aes(x=INCOME20), bins=50)
ggplot(delta_vulnerable_shp3) + geom_histogram(aes(x=NOENG),    bins=50)
ggplot(delta_vulnerable_shp3) + geom_histogram(aes(x=NOHISCH),  bins=50)

# log(data)
ggplot(delta_vulnerable_shp3) + geom_histogram(aes(x=log(deltaHg)),  bins=50)
ggplot(delta_vulnerable_shp3) + geom_histogram(aes(x=log(POVERTY)),  bins=50)
ggplot(delta_vulnerable_shp3) + geom_histogram(aes(x=log(FISH)),     bins=50)
ggplot(delta_vulnerable_shp3) + geom_histogram(aes(x=log(INCOME20)), bins=50)
ggplot(delta_vulnerable_shp3) + geom_histogram(aes(x=log(NOENG)),    bins=50)
ggplot(delta_vulnerable_shp3) + geom_histogram(aes(x=log(NOHISCH)),  bins=50)

```

Transform values into normal distribution (save log()).
```{r}
## Create spatial lag eqn
delta_vulnerable_shp3 <- delta_vulnerable_shp3 %>% st_as_sf() %>% mutate(
  logHg       = log(deltaHg),
  logPOVERTY  = log(POVERTY),
  logFISH     = log(FISH),
  logPOVERTY  = log(POVERTY),
  logINCOME20 = log(INCOME20),
  logNOHISCH  = log(NOHISCH),
  logNOENG    = log(NOENG)
)

delta_vulnerable_shp3 <-
  delta_vulnerable_shp3[is.finite(rowSums(delta_vulnerable_shp3[, c('logHg', 'logFISH', 'logINCOME20', 'logPOVERTY', 'logNOENG', 'logNOHISCH')] %>% st_drop_geometry())), ]

delta_vulnerable_shp3

```

Perform spatial regression.
```{r}
logHg_slm <- Spmodel_log(delta_vulnerable_shp3)

# Save model
slm <- logHg_slm$slm
# Save neighborhood matrix
Nb  <- logHg_slm$Nb

summary(slm)

```

Calculate coefficients.
```{r}
coef(slm)
confint(slm)

TabSLM(slm)

```

Check Moran's I - Did it get smaller?
```{r}
# Derive the residuals from the regression. Need to handle those missed values
slResiduals <- rep(0, length(delta_vulnerable_shp3$deltaHg))
resIndex    <- slm$residuals %>% names() %>% as.integer();
slResiduals[resIndex] <- slm$residuals

MoranI_orig <- Nb %>%
  spdep::moran.test(delta_vulnerable_shp3$deltaHg, ., zero.policy = TRUE)

MoranI_slm <- Nb %>%
  spdep::moran.test(slResiduals, ., zero.policy = TRUE)

MoranI_orig
MoranI_slm

```

Calculate Local Indicators of Spatial Association (LISA) - Identify clusters of high or low values as well as outliers that are surrounded by opposite values.
```{r}
# USA
a <- plotCOType(delta_vulnerable_shp3,
                'deltaHg',
                'deltaHg = (Hg2010 - Hg2020) / Hg2010') + geom_sf(data = usa_shp, alpha = 0) #+ theme(legend.position = "none")
b <- plotCOType(delta_vulnerable_shp3,
                'FISH', "Median High Fish Consumers") + geom_sf(data = usa_shp, alpha = 0) + theme(legend.position = "none")
c <- plotCOType(delta_vulnerable_shp3,
                'INCOME20', "Income <$20k") + geom_sf(data = usa_shp, alpha = 0) + theme(legend.position = "none")
d <- plotCOType(delta_vulnerable_shp3,
                'POVERTY', "PIR <2") + geom_sf(data = usa_shp, alpha = 0) + theme(legend.position = "none")
e <- plotCOType(delta_vulnerable_shp3,
                'NOHISCH', "<High School Education") + geom_sf(data = usa_shp, alpha = 0) + theme(legend.position = "none")
f <- plotCOType(delta_vulnerable_shp3,
                'NOENG',
                "Limited English Speaking Household") + geom_sf(data = usa_shp, alpha = 0) + theme(legend.position = "none")

a
b
c
d
e
f

# plot_grid(a,b,c,d,e,f)


```








